descriptive_Clusters

### 2020-10-21 src/descriptive_Clusters.Rmd
More writing...

As people came to rely more on their broadband services, it is unsurprising that both the amount and rate of testing increased in 2020 during the lockdown compared to the same period in 2019. This increase was greatest in cluster 2, where 56% more tests were run compared to the same period in 2019. Cluster 2 also has the most local authorities, 148, which are home to more than 28.5 million people, so the sum of tests run in all these authorities is much higher than in all the other clusters, both in 2019 and 2020. Cluster 1 in comparison, not only has has the lowest  frequency of testing per capita in both 2019 and 2020, it also shows the lowest rate of increase in testing between the two time periods at just 33%. Meanwhile, cluster 5 has the fewest local authorities, at 28, but the second highest total number of tests in both periods and the second largest increase in testing between 2019 and 2020, at 52%. It also has the highest number of tests per capita in both periods. A possible explanation is that cluster 5, with a population of just under 12 million contains some of the most populous local authorities in the country, including the cities of Birmingham, Manchester, Leeds, Sheffield, Glasgow and Edinburgh. However, this cluster also includes large rural counties, such as Cornwall and Northumberland. Are urban dwellers or rural households more likely to test their broadband speeds?

@@

The increase in daytime compared to evening testing is particularly noticeable for cluster 2, but the change is less pronounced for cluster 5. Indeed, in an auxiliary regression, the probability of additional tests per person run between 9.00 and 11.00 was significantly higher for those living in cluster 2 than in cluster 5, whilst the probability was significantly lower in cluster 1. ?? perhaps because of the diversity of urban and rural populations represented in cluster 5 ?? 
#re-base auxiliary regression so reference case is either 1 or 3 - we wanted to talk most about 5?
The graphs also show that Cluster 1, with the lowest increase in testing, falls from the middle position in terms of total number of tests to the second from bottom.

For upload speeds, 345 of 382 local authorities, or over 62 million people, fall into cluster 6 or cluster 9. Graph [.] shows that both of these clusters have relatively similar temporal profiles, with relatively stable speeds. However, the upload speeds at all times for cluster 9 are substantially higher than for cluster 6. This difference is reflected not only in the mean speeds for these clusters for the whole sample, but also the mean upload speeds in evening peak period from 19:00-20:59 and the morning peak from 9:00-10:59. The 115 Local Authority areas in cluster 9 are mostly in urban or suburban areas, including 13 London Boroughs (of 32), 8 of the 10 local authorities of Greater Manchester, 5 of the 7 constituent authorities of the West Midlands Combined Authority, as well as cities like Glasgow, Leicester, Nottingham, Sheffield, and the Portsmouth and Southampton conurbation. There are also some notable medium-sized cities, including Aberdeen, Cardiff, Oxford, Milton Keynes, and York, as well as many suburban districts from the South East of England to South Tyneside.

Meanwhile, the 230 local authorities in cluster 6 still include major urban areas, such as Bristol, Liverpool and Leeds, and many suburban areas, but also include some of the most rural areas in the country. Slower speeds could be seen as a typical level of service in such areas, and indeed, upload speeds in cluster 6 are only 4% slower in the morning than in the evening, compared to 12-19% slower for the almost 3 million people living in five of the smaller clusters. And yet, auxiliary regressions show that those living in cluster 6 had the highest probability of testing their broadband between 9:00 and 11:00 of any of the 10 upload speed clusters. Cluster 6 also experienced the joint highest ratio of standard deviation to mean across the time period under assessment. This suggests that although the time profile is relatively flat in graph [.], the experience is one of speeds that fluctuate from a lower mean, and therefore might more often impact on the activities users are doing.

Cluster 1, with 10 local authorities that are home to over 1 million people, is also interesting, as it has the second slowest speeds in the morning compared to the evening 'peaks' and the second highest ratio of standard deviation to mean. Indeed, the time profile in graph [.] shows upload speeds in cluster 1 are on average lower at certain times of day during the study period than any other cluster. Yet the auxiliary regression suggests that there are not many tests being run during the am peak in cluster 1. This may be because there are fewer people working at home, anxiously checking their broadband than in most other clusters. Indeed, the auxiliary regressions indicate that cluster 1 has the highest job density or proportion of jobs to working-age population, which is likely to due to the presence of Westminster, central London, cluster 1's most populous local authority. Westminster not only has more workplaces than residents, but it is reasonable to presume that many who would normally work in Westminster, but be able to work from home during lockdown are likely to live outside central London and not be subject to the fluctuating speeds there. Workplaces, meanwhile, some of which would still be open in a place like Westminster, could be running programmes that cause the slowdown and variation, but would be more likely to have their own in-house diagnostics, rather than using a service like Speedchecker Ltd.

Furthermore, whilst cluster 1 has the second highest number of established businesses per capita, it is far behind the cluster with the most business per capita, cluster 6. This suggests that cluster 1 has more large employers than cluster 6, which fits with the likelihood that those working from home whose large employer is based in cluster 1 are much less likely than those in cluster 6 to also live in a local authority in cluster 1 (or at least Westminster, which probably skews the sample). Meanwhile, cluster 9 has the lowest number of established businesses per inhabitant. Considering how many central urban areas fall into cluster 9, this suggests an even greater dominance of large employers. Yet the higher upload speeds in cluster 9 may mean that people working from home in these local authorities noticed fewer problems. There is also the least difference between morning and evening upload speeds for cluster 9, at 1% slower in the morning. However, that is not to say that cluster 9 did not include many people telecommuting during lockdown.

Indeed, the only cluster where upload speeds were slower in the evening than in the morning was cluster 5, made up of 2 local authorities with less than 200,000 people: Three Rivers, a suburban district north of London, and Fylde, a seaside suburb of Blackpool. The auxiliary regressions show that compared to these two places, all the other clusters had a lower percentage of working people in managerial, professional and administrative jobs.???

This is an indication of the reliability of broadband services in places in both clusters, and suggests that upload speeds have generally been robust throughout the study period. The exceptions can be found in seven of the eight much smaller clusters including 35 local authority districts, where AM peak upload speeds are between 6% and 18% slower than PM peak upload speeds, although the mean speeds for each cluster are higher than cluster 6. The temporal profiles on Graph [.] for the three of those clusters with more than 5 local authorities also shows that speeds are far from dependable in these clusters. Indeed, in 25 local authorities with a combined population of almost 3 million, speeds are 13% or more slower in the morning than in the evening. Included in this latter group are central London borough of Westminster and the London Borough of Newham, rural authorities like Eden and West Devon, and small cities like Dundee and Carlisle.

>>>>

As people came to rely more on their broadband services, it is unsurprising that both the amount and rate of testing increased in 2020 during the lockdown compared to the same period in 2019. This increase was greatest in cluster 2, where 56% more tests were run compared to the same period in 2019. Cluster 2 also has the most local authorities, 148, which are home to more than 28.5 million people, so the sum of tests run in all these authorities is much higher than in all the other clusters, both in 2019 and 2020. Cluster 1 in comparison, not only has has the lowest  frequency of testing per capita in both 2019 and 2020, it also shows the lowest rate of increase in testing between the two time periods at just 33%.

Meanwhile, cluster 5 has the fewest local authorities, at 28, but the second highest total number of tests in both periods and the second largest increase in testing between 2019 and 2020, at 52%. It also has the highest number of tests per capita in both periods. A possible explanation is that cluster 5, with a population of just under 12 million contains some of the most populous local authorities in the country, including the cities of Birmingham, Manchester, Leeds, Sheffield, Glasgow and Edinburgh. However, this cluster also includes large rural counties, such as Cornwall and Northumberland. Are urban dwellers or rural households more likely to test their broadband speeds? Auxiliary regressions suggest that cluster 5 also had the highest earnings of the clusters, but was least likely to have a Virgin Media connection.

@@

The increase in daytime compared to evening testing is particularly noticeable for cluster 2, but the change is less pronounced for cluster 5. Indeed, in an auxiliary regression, the probability of additional tests per person run between 9.00 and 11.00 was significantly higher for those living in cluster 2 than in cluster 5, but both were higher than the other three clusters. ?? perhaps because of the diversity of urban and rural populations represented in cluster 5 ?? Cluster 5 mostly same as cluster 6 upload, few cluster 9, but same for cluster 2?
#re-base auxiliary regression so reference case is either 1 or 3 - we wanted to talk most about 5?
The graphs also show that Cluster 1, with the lowest increase in testing, falls from the middle position in terms of total number of tests to the second from bottom.

For upload speeds, 345 of 382 local authorities, or over 62 million people, fall into cluster 6 or cluster 9. Graph [.] shows that both of these clusters have relatively similar temporal profiles, with relatively stable speeds. However, the upload speeds at all times for cluster 9 are substantially higher than for cluster 6. This difference is reflected not only in the mean speeds for these clusters for the whole sample, but also the mean upload speeds in the morning peak from 9:00-10:59, as well as the evening peak period from 19:00-20:59. The quality of broadband service is thus much better in the 115 Local Authority areas in cluster 9, which are mostly in urban or suburban areas. These include 13 London Boroughs (of 32), 8 of the 10 local authorities of Greater Manchester, 5 of the 7 constituent authorities of the West Midlands Combined Authority, as well as cities like Glasgow, Leicester, Nottingham, Sheffield, and the Portsmouth and Southampton conurbation. There are also some notable medium-sized cities, including Aberdeen, Cardiff, Oxford, Milton Keynes, and York, and many suburban districts from the South East of England to South Tyneside. Auxiliary regressions indicate that the speed tests in these authorities are also more likely to have been run on services provided by Virgin Media, suggesting they are in the half of the country with the most lucrative ICT market, which originally attracted the cable TV provider (OfCom...).

Meanwhile, the 230 local authorities in cluster 6 still include major urban areas, such as Bristol, Liverpool and Leeds, and many suburban areas, but also include some of the most rural areas in the country. For example, although auxiliary regressions show that Cluster 6 local authorities are more likely to be in the South of the UK than Cluster 9, the cluster notably includes Southern rural districts from Cornwall to North Norfolk. Slower speeds could reflect the lower quality of service in rural areas compared to urban and suburban areas. The reliability of service in these areas is also in question. For although upload speeds in cluster 6 are only 4% slower in the morning than in the evening, compared to 12-19% slower for the almost 3 million people living in five of the smaller clusters, auxiliary regressions show that those living in cluster 6 had the highest probability of testing their broadband between 9:00 and 11:00 of any of the 10 upload speed clusters. Cluster 6 also experienced the joint highest ratio of standard deviation to mean across the time period under assessment. This suggests that although the time profile is relatively flat in graph [.], the experience is one of speeds that fluctuate from a lower mean, and therefore might more often impact on the activities users are doing.

In comparison, there is negligible difference between morning and evening upload speeds for cluster 9, at 1% slower in the morning. This suggests a high level of reliability of service, a conclusion supported by the finding in the auxiliary regressions that those in cluster 9 ran the fewest speed tests per person during the morning period of any cluster. Now this may be an indication of fewer people working from home, less contention, and less resultant frustration. Cluster 9 comprises many central urban areas and has the lowest number of established businesses per inhabitant, which could be interpreted as a dominance of large employers. However, the job density is lower than cluster 6, meaning there are not as many jobs per resident in these areas. As the cluster includes many suburban areas too, which may be largely residential, could the quality and reliability of internet service be reinforcing patterns of telecommuting by those in wealthier suburbs who can work from home? Earnings in cluster 9 are second highest of all the clusters, with only Cluster 2 (comprising just North East Lincolnshire and East Lothian, population 265k) earning more per person.

Meanwhile, Cluster 1, with 10 local authorities that are home to over 1 million people has the second slowest speeds in the morning compared to the evening 'peaks' and the second highest ratio of standard deviation to mean. Indeed, the time profile in graph [.] shows upload speeds in cluster 1 are on average lower at certain times of day during the study period than any other cluster. Yet the auxiliary regression suggests that there are not many tests being run during the am peak in cluster 1. This may be because there are fewer people working at home checking their broadband than in most other clusters. Indeed, the auxiliary regressions indicate that cluster 1 has the highest job density or proportion of jobs to working-age population, which is likely to due to the presence of Westminster, central London, cluster 1's most populous local authority. Westminster not only has more workplaces than residents, but it is reasonable to presume that many who would normally work in Westminster, but be able to work from home during lockdown are likely to live outside central London and not be subject to the fluctuating speeds there. Workplaces, meanwhile, some of which would still have been open, could be running programmes that cause the slowdown and variation, but would be more likely to have their own in-house diagnostics, rather than using a service like Speedchecker Ltd.

Indeed, the only cluster where upload speeds were slower in the evening than in the morning was cluster 5, made up of 2 local authorities with less than 200,000 people: Three Rivers, a suburban district north of London, and Fylde, a seaside suburb of Blackpool. However, these are likely to be outliers and may not have many tests from which the clusters are calcualted. **Take out smaller clusters!?** The auxiliary regressions show that compared to these two places, all the other clusters had a lower percentage of working people in managerial, professional and administrative jobs.???

This is an indication of the reliability **or maybe not - 5% slowdown from low base can be noticeable, and sd of cluster 6 bigger** of broadband services in places in both clusters, and suggests that upload speeds have generally been robust throughout the study period. The exceptions can be found in seven of the eight much smaller clusters including 35 local authority districts, where AM peak upload speeds are between 6% and 18% slower than PM peak upload speeds, although the mean speeds for each cluster are higher than cluster 6. The temporal profiles on Graph [.] for the three of those clusters with more than 5 local authorities also shows that speeds are far from dependable in these clusters. Indeed, in 25 local authorities with a combined population of almost 3 million, speeds are 13% or more slower in the morning than in the evening. Included in this latter group are central London borough of Westminster and the London Borough of Newham, rural authorities like Eden and West Devon, and small cities like Dundee and Carlisle.
